
---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, comment="#>")
#knitr::opts_chunk$set(message=FALSE, comment="#>", fig.path = "docs/README_files/figure-gfm/")
library(tidyverse)
devtools::load_all(".")
```

# CaDrA

<!-- badges: start -->
  [![R-CMD-check](https://github.com/montilab/CaDrA/workflows/R-CMD-check/badge.svg)](https://github.com/montilab/CaDrA/actions)
<!-- badges: end -->
  
Candidate Drivers Analysis: Multi-Omic Search for Candidate Drivers of Functional Signatures

CaDrA is an R package that supports a heuristic search framework aimed towards the identification of candidate drivers of oncogenic activity. Given a binary genomic dataset (where the rows are 1/0 vectors indicating the presence/absence of genomic features such as somatic mutations or copy number alteration events), together with an associated sample ranking (where the samples are ranked by a certain phenotypic readout of interest such as protein expression, pathway activity etc.), CaDrA implements a step-wise search algorithm to determine a set of features that, together (based on their occurence union or 'logical OR'), is most-associated with the observed ranking, making it useful for finding mutually exclusive or largely non-overlapping anomalies that can lead to the same pathway phenotype.

For more information, please see the associated manuscript [Kartha et al. (2019)](https://www.frontiersin.org/articles/10.3389/fgene.2019.00121/full)

## (1) Installation

You can install the development version of CaDrA from GitHub (**Recommended**)

```r
library(devtools)
devtools::install_github("montilab/CaDrA", ref="dev")
```

## (2) Quickstart

```{r, eval=FALSE} 
library(CaDrA)
```

### Test run code on simulated data

```{r, fig.width=8, fig.align='center'}

# Load pre-computed Top-N list generated for sim.ES dataset
data(topn.list)

# Plot the results from a top-N evaluation by passing the resulting ESet from a specific run
# To find the combination of features that had the best score
best_meta <- topn_best(topn_list = topn.list) 

# Now we can plot this set of features
meta_plot(topn_best_list = best_meta)

```

```{r, fig.width=8, fig.align='center'}

# Get top N plot
topn_plot(topn_list = topn.list)

```

### Running on actual data

In the example below, the `CCLE_MUT_SCNA` expression set object has 17723 rows (features) and 82 columns (samples). The other variable is called `CTNBB1_reporter` which contains continuous measures of a targeted profile. The **CTNBB1_reporter** is what dictates how CaDrA will search for grouped meta-features.

```{r, fig.width=8, fig.align='center'}

# Load binary feature data object
data(CCLE_MUT_SCNA)

# Load a targeted profile object
data(CTNBB1_reporter)

# Number of top starting seed features to test and evaluate over  
top_N <- 7

# Metric used for candidate search
# Either ks or wilcox or revealer or custom function is supported
method <- "ks"

topn_l <- topn_eval(
  ES = CCLE_MUT_SCNA, 
  input_score =  CTNBB1_reporter,
  method = method,
  alternative = "less",
  metric = "pval",
  top_N = top_N,
  search_method = "both",
  max_size = 7,
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # Set best_score_only = FALSE will return both the eset and best scores
)

# Now we can fetch the ESet and feature that corresponded to the best score over the top N search
topn_best_meta <- topn_best(topn_l)

# Visualize best meta-features result
meta_plot(topn_best_list = topn_best_meta)

# You can also evaluate how robust the results are depending on which seed feature you started with
topn_plot(topn_l) 

```

