---
title: "Scoring Functions"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Scoring Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, collapse = TRUE, comment="")
library(SummarizedExperiment)
library(devtools)
load_all()
```

The **CaDrA** package currently provides four scoring functions to search for subsets of genomic features that are likely associated with a specific outcome of interest (e.g., protein expression, pathway activity, etc.)

1. Kolmogorov-Smirnov Method (`ks`)
2. Conditional Mutual Information Method (`knnmi`)
3. Wilcoxon Rank-Sum Method (`wilcox`)
4. Custom - An User's Provided Scoring Function (`custom`)

Below, we run `candidate_search()` over the top 3 starting features using each of the four scoring functions described above.

**Important Note:**

- The legacy function `topn_eval()` is equivalent to the recommended `candidate_search()` function

# Load packages

```r
library(CaDrA)
```

# Load required datasets

1. A `binary features matrix` also known as `Feature Set` (such as somatic mutations, copy number alterations, chromosomal translocations, etc.) The 1/0 row vectors indicate the presence/absence of ‘omics’ features in the samples. The `Feature Set` must be an object of class **SummarizedExperiment** from **SummarizedExperiment** package) 
2. A vector of continuous scores (or `input_score`) representing a functional response of interest (such as protein expression, pathway activity, etc.)

```{r load.data}

# Load pre-computed feature set
data(sim_FS)
 
# Load pre-computed input scores
data(sim_Scores)

```

# Kolmogorov-Smirnov scoring method 

See `?ks_rowscore` for more details

```{r ks.method}

ks_topn_l <- CaDrA::candidate_search(
  FS = sim_FS,
  input_score = sim_Scores,
  method = "ks_pval",          # Use Kolmogorow-Smirnow scoring function 
  weights = NULL,              # If weights is provided, perform a weighted-KS test
  alternative = "less",        # Use one-sided hypothesis testing
  search_method = "both",      # Apply both forward and backward search
  top_N = 3,                   # Evaluate top 3 starting points for the search
  max_size = 7,                # Allow at most 7 features in meta-feature matrix
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # Return meta-feature, its observed input scores and corresponding best score
)

# Now we can fetch the feature set of top N features that corresponded to the best scores over the top N search
ks_topn_best_meta <- topn_best(ks_topn_l)

# Visualize best meta-feature result
meta_plot(topn_best_list = ks_topn_best_meta)

```

# Wilcoxon Rank-Sum scoring method 

See `?wilcox_rowscore` for more details

```{r wilcox.method}

wilcox_topn_l <- CaDrA::candidate_search(
  FS = sim_FS,
  input_score = sim_Scores,
  method = "wilcox_pval",      # Use Wilcoxon Rank-Sum scoring function
  alternative = "less",        # Use one-sided hypothesis testing
  search_method = "both",      # Apply both forward and backward search
  top_N = 3,                   # Evaluate top 3 starting points for the search
  max_size = 7,                # Allow at most 7 features in meta-feature matrix
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # Return meta-feature, its observed input scores and corresponding best score
)

# Now we can fetch the feature set of top N feature that corresponded to the best scores over the top N search
wilcox_topn_best_meta <- topn_best(topn_list = wilcox_topn_l)

# Visualize best meta-feature result
meta_plot(topn_best_list = wilcox_topn_best_meta)

```

# Conditional Mutual Information scoring method 

See `?knnmi_rowscore` for more details

```{r knnmi.method}

knnmi_topn_l <- CaDrA::candidate_search(
  FS = sim_FS,
  input_score = sim_Scores,
  method = "knnmi",            # Use knnmi CMI scoring function
  search_method = "both",      # Apply both forward and backward search
  top_N = 3,                   # Evaluate top 3 starting points for the search
  max_size = 7,                # Allow at most 7 features in meta-feature matrix
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # Return meta-feature, its observed input scores and corresponding best score
)

# Now we can fetch the ESet of top feature that corresponded to the best scores over the top N search
knnmi_topn_best_meta <- topn_best(topn_list = knnmi_topn_l)

# Visualize best meta-feature result
meta_plot(topn_best_list = knnmi_topn_best_meta)

```

# Custom - An user's provided scoring method

See `?custom_rowscore` for more details

```{r custom.method}

# A customized function using ks-test function
customized_ks_rowscore <- function(FS, input_score, meta_feature=NULL, alternative="less"){
   
   # Check if meta_feature is provided
   if(!is.null(meta_feature)){
     # Getting the position of the known meta features
     locs <- match(meta_feature, row.names(FS))
     
     # Taking the union across the known meta features
     if(length(meta_feature) > 1) {
       meta_vector <- as.numeric(ifelse(colSums(FS[meta_feature,]) == 0, 0, 1))
     }else{
       meta_vector <- as.numeric(FS[meta_feature,])
     }
      
     # Remove the meta features from the binary feature matrix
     # and taking logical OR btw the remaining features with the meta vector
     locs <- match(meta_feature, row.names(FS))
     FS <- base::sweep(FS[-locs, , drop=FALSE], 2, meta_vector, `|`)*1
      
     # Check if there are any features that are all 1s generated from
     # taking the union between the matrix
     # We cannot compute statistics for such features and thus they need
     # to be filtered out
     if(any(rowSums(FS) == ncol(FS))){
       warning("Features with all 1s generated from taking the matrix union ",
               "will be removed before progressing...\n")
       FS <- FS[rowSums(FS) != ncol(FS), , drop=FALSE]
     }
   }
    
   # KS is a ranked-based method
   # So we need to sort input_score from highest to lowest values
   input_score <- sort(input_score, decreasing=TRUE)
    
   # Re-order the matrix based on the order of input_score
   FS <- FS[, names(input_score), drop=FALSE]  
   
   # Compute the scores using the KS method
   ks <- apply(FS, 1, function(r){ 
     x = input_score[which(r==1)]; 
     y = input_score[which(r==0)];
     res <- ks.test(x, y, alternative=alternative)
     return(c(res$statistic, res$p.value))
   })
   
   # Obtain score statistics
   stat <- ks[1,]
   
   # Obtain p-values and change values of 0 to the machine lowest value 
   # to avoid taking -log(0)
   pval <- ks[2,]
   pval[which(pval == 0)] <- .Machine$double.xmin
   
   # Compute the -log(pval)
   # Make sure scores has names that match the row names of FS object
   scores <- -log(pval)
   names(scores) <- rownames(FS)
   
   return(scores)
   
}

# Search for best features using a custom-defined function
custom_topn_l <- CaDrA::candidate_search(
  FS = SummarizedExperiment::assay(sim_FS),
  input_score = sim_Scores,
  method = "custom",                                # Use custom scoring function
  custom_function = customized_ks_rowscore,         # Use a customized scoring function
  custom_parameters = NULL,                         # Additional parameters to pass to custom_function
  search_method = "both",                           # Apply both forward and backward search
  top_N = 3,                                        # Evaluate top 3 starting points for the search
  max_size = 7,                                     # Allow at most 7 features in meta-feature matrix
  do_plot = FALSE,                                  # We will plot it AFTER finding the best hits
  best_score_only = FALSE                           # Return meta-feature, its observed input scores and corresponding best score
)

# Now we can fetch the feature set of top N feature that corresponded to the best scores over the top N search
custom_topn_best_meta <- topn_best(topn_list = custom_topn_l)

# Visualize best meta-feature result
meta_plot(topn_best_list = custom_topn_best_meta)

```

# SessionInfo

```{r RsessionInfo}
sessionInfo()
```


