---
title: "Scoring Functions"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Scoring Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, collapse = TRUE, comment="", 
                      fig.align="center")
library(Biobase)
devtools::load_all(".")
```

In **CaDrA** package, there are four scoring functions that were 
implemented to search for relevant subsets of genomic features that are 
likely associated with a specific outcome of interest 
(i.e. protein expression, pathway activity, etc.)

1. Kolmogorov-Smirnov Method (`ks`)
2. Conditional Mutual Information Method (`revealer`)
3. Wilcoxon Rank-Sum Method (`wilcox`)
4. Custom - User Provided Scoring Function (`custom`)

**Important Note:**

- To search for best meta-feature over top N starting seed features, see `?candidate_search()` or `?topn_eval()`
- Both `candidate_search()` and `topn_eval()` are basically the same function 
as we previously combined functionality of `topn_eval()` 
into `candidate_search()` as one unified function

Below, we are evaluating the candidate search over top 7 starting seed 
features using four different scoring methods mentioned above.

# Load packages

```r
library(CaDrA)
library(Biobase)
```

# Load required datasets

1. A `binary features matrix` also known as `Feature Set` (such as 
somatic mutations, copy number alterations, chromosomal translocations, etc.) 
The 1/0 vectors indicating the presence/absence of ‘omics’ features 
in the samples. The `Feature Set` must be an object of class **ExpressionSet** 
from **Biobase** package) 
2. A vector of continuous scores (or `input_score`) represents 
a functional response of interest (such as protein expression, 
pathway activity, etc.)

```{r load.data}

## Read in BRCA GISTIC+Mutation ESet object
data(BRCA_GISTIC_MUT_SIG)
eset_mut_scna <- BRCA_GISTIC_MUT_SIG

## Read in input score of TAZYAP activity
data(TAZYAP_BRCA_ACTIVITY)
input_scores <- TAZYAP_BRCA_ACTIVITY

## Samples to keep based on the overlap between the two inputs
overlap <- intersect(names(input_scores), Biobase::sampleNames(eset_mut_scna))
eset_mut_scna <- eset_mut_scna[,overlap]
input_scores <- input_scores[overlap]

## Binarize ES to only have 0's and 1's
exprs(eset_mut_scna)[exprs(eset_mut_scna) > 1] <- 1.0

## Pre-filter ESet based on occurrence frequency
eset_mut_scna_flt <- CaDrA::prefilter_data(
  ES = eset_mut_scna,
  max.cutoff = 0.6,  # max event frequency (60%)
  min.cutoff = 0.03  # min event frequency (3%)
) 

# Define required variables
ES = eset_mut_scna_flt
input_score = input_scores

```

# Kolmogorov-Smirnov scoring method 

See `?ks_genescore_mat` for more details

```{r ks.method}

ks_topn_l <- CaDrA::candidate_search(
  ES = ES,
  input_score = input_score,
  method = "ks",               # Use Kolmogorow-Smirnow scoring function 
  weights = NULL,              # If weights is provided, it will be used 
                               # to perform a weighted-KS test
  alternative = "less",        # Use one-sided p-value
  metric = "pval",             # Use p-value to search for best feature
  search_method = "both",      # Apply both forward and backward search
  top_N = 7,                   # Evaluate top 7 starting points for the search
  max_size = 10,               # Set max_size = 10 to account for 
                               # 10 left-skewed features in sim.ES
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # If best_score_only = FALSE, it will 
                               # return meta-feature Eset, 
                               # its observed input scores and corresponding 
                               # best score
)

# Now we can fetch the ESet of top feature that corresponded to the best 
# scores over the top N search
ks_topn_best_meta <- topn_best(ks_topn_l)

# Visualize best meta-feature result
meta_plot(topn_best_list = ks_topn_best_meta)

```

# Wilcoxon Rank-Sum scoring method 

See `?wilcox_genescore_mat` for more details

```{r wilcox.method}

wilcox_topn_l <- CaDrA::candidate_search(
  ES = ES,
  input_score = input_score,
  method = "wilcox",           # Use Wilcoxon Rank-Sum scoring function
  alternative = "less",        # Use one-sided p-value
  metric = "pval",             # Use p-value to search for best feature
  search_method = "both",      # Apply both forward and backward search
  top_N = 7,                   # Evaluate top 7 starting points for the search
  max_size = 10,               # Set max_size = 10 to account for 
                               # 10 left-skewed features in sim.ES
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # If best_score_only = FALSE, it will return 
                               # meta-feature Eset, 
                               # its observed input scores and corresponding 
                               # best score
)

# Now we can fetch the ESet of top feature that corresponded to 
# the best scores over the top N search
wilcox_topn_best_meta <- topn_best(wilcox_topn_l)

# Visualize best meta-feature result
meta_plot(topn_best_list = wilcox_topn_best_meta)

```

# Conditional Mutual Information scoring method 

See `?revealer_genescore_mat` for more details

```{r revealer.method}

revealer_topn_l <- CaDrA::candidate_search(
  ES = ES,
  input_score = input_score,
  method = "revealer",         # Use REVEALER's CMI scoring function
  metric = "stat",             # Use score statistics to search for best feature
  search_method = "both",      # Apply both forward and backward search
  top_N = 7,                   # Evaluate top 7 starting points for the search
  max_size = 10,               # Set max_size = 10 to account for 
                               # 10 left-skewed features in sim.ES
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # If best_score_only = FALSE, it will return 
                               # meta-feature Eset, 
                               # its observed input scores and corresponding 
                               # best score
)

# Now we can fetch the ESet of top feature that corresponded to the 
# best scores over the top N search
revealer_topn_best_meta <- topn_best(revealer_topn_l)

# Visualize best meta-feature result
meta_plot(topn_best_list = revealer_topn_best_meta)

```

# Custom - user provided scoring method

See `?custom_genescore_mat` for more details

```{r custom.method}

# A customized function using ks-test function
customized_genescore_mat <- function(mat, input_score, alternative){
  result <- 1:nrow(mat) %>% 
    purrr::map_dfr(
      function(r){ 
        feature = mat[r,];
        x = input_score[which(feature==1)]; y = input_score[which(feature==0)];
        res <- ks.test(x, y, alternative=alternative)
        return(data.frame(score=res$statistic, p_value=res$p.value))
   })
}

# Search for best features using a custom-defined function
custom_topn_l <- CaDrA::candidate_search(
  ES = ES,
  input_score = input_score,
  custom_function = customized_genescore_mat,# Use a customized scoring function
  # Additional parameters to pass to custom_function
  custom_parameters = list(alternative = "less"),   
  metric = "pval",          # Use pval to search for best feature
  search_method = "both",   # Apply both forward and backward search
  top_N = 7,                # Evaluate top 7 starting points for the search
  max_size = 10,            # Set max_size = 10 to account for 
                            # 10 left-skewed features in sim.ES
  do_plot = FALSE,          # We will plot it AFTER finding the best hits
  best_score_only = FALSE   # If best_score_only = FALSE, it will return 
                            # meta-feature Eset,  its observed input scores 
                            # and corresponding best score
)

# Now we can fetch the ESet of top feature that corresponded to the 
# best scores over the top N search
custom_topn_best_meta <- topn_best(custom_topn_l)

# Visualize best meta-feature result
meta_plot(topn_best_list = custom_topn_best_meta)

```


## R Session Info
```{r RsessionInfo}
sessionInfo()
```