---
title: "Scoring Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scoring Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(comment="", cache=FALSE, fig.width=8, fig.height=8, fig.align="center")
library(Biobase)
devtools::load_all(".")
```

In **CaDrA** package, there are four scoring functions that were implemented to search for relevant subsets of genomic features that are likely associated with a specific outcome of interest (i.e. `input_score`):

1. Kolmogorov-Smirnov Method (`ks`)
2. Conditional Mutual Information Method (`revealer`)
3. Wilcoxon Rank-Sum Method (`wilcox`)
4. Custom-Defined Scoring Function (`custom`)

**Important Note:**

- To search for best meta-features over top N starting seed features, see `?topn_eval()`
- To evaluate with a specific starting seed feature of interest, see `?candidate_search()`

Here, we test and evaluate top 7 starting seed features using four scoring methods mentioned above.

# Load packages

```r
library(CaDrA)
library(Biobase)
```

# Load required datasets

1. A `binary features matrix` (such as somatic mutations, copy number alterations, chromosomal translocations, etc.). The 1/0 vectors indicating the presence/absence of ‘omics’ features in the samples. The `features` matrix must be an object of class **ExpressionSet** from **Biobase** package) 
2. A vector of continuous scores (or `input_score`) represents a functional response of interest (such as protein expression, pathway activity, etc.)

```{r load.data}

# Load pre-simulated dataset
# See ?sim.ES for more information
data(sim.ES)
ES = sim.ES

# Generate a vector of sample-specific scores representing a target profile of interest
set.seed(123)                             # set seed
input_score = rnorm(n = ncol(sim.ES))     # generate random scores
names(input_score) = colnames(sim.ES)     # label input_score with names similar to colnames of sim.ES


```

## 1. Kolmogorov-Smirnov scoring method 

See `?ks_genescore_mat` for more details

```{r ks.method}

ks_topn_l <- CaDrA::topn_eval(
  ES = ES,
  input_score = input_score,
  method = "ks",               # Use Kolmogorow-Smirnow Scoring function 
  weights = NULL,              # If weights is provided, it will be used to perform a weighted-KS test
  alternative = "less",        # one-sided p-value
  metric = "stat",             # Use p-value to score features
  top_N = 7,                   # Evaluate top 7 starting points for the search
  max_size = 10,               # Set max_size = 10 to account for 10 left-skewed features in sim.ES
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # If best_score_only = FALSE, it will return ESet, its corresponding
                               # best score, and input_score for over 7 feature searches
)

# Now we can fetch the ESet and feature that corresponded to the best score over the top N search
ks_topn_best_meta <- topn_best(ks_topn_l)

# Visualize best meta-features result
meta_plot(topn_best_list = ks_topn_best_meta)

```

## 2. Wilcoxon Rank-Sum scoring method 

See `?wilcox_genescore_mat` for more details

```{r wilcox.method}

wilcox_topn_l <- CaDrA::topn_eval(
  ES = ES,
  input_score = input_score,
  method = "wilcox",           # Use Wilcoxon Rank-Sum scoring function
  alternative = "less",        # one-sided p-value
  metric = "stat",             # Use p-value to score features
  top_N = 7,                   # Evaluate top 7 starting points for the search
  max_size = 10,               # Set max_size = 10 to account for 10 left-skewed features in sim.ES
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # If best_score_only = FALSE, it will return ESet, its corresponding
                               # best score, and input_score for over 7 feature searches
)

# Now we can fetch the ESet and feature that corresponded to the best score over the top N search
wilcox_topn_best_meta <- topn_best(wilcox_topn_l)

# Visualize best meta-features result
meta_plot(topn_best_list = wilcox_topn_best_meta)

```

## 3. Conditional Mutual Information scoring method 

See `?revealer_genescore_mat` for more details

```{r revealer.method}

revealer_topn_l <- CaDrA::topn_eval(
  ES = ES,
  input_score = input_score,
  method = "revealer",         # Use REVEALER's CMI scoring function
  target_match = "positive",   # A direction of target matching. "positive" for higher values
  metric = "stat",             # Use score statistics to score features
  top_N = 7,                   # Evaluate top 7 starting points for the search
  max_size = 10,               # Set max_size = 10 to account for 10 left-skewed features in sim.ES
  do_plot = FALSE,             # We will plot it AFTER finding the best hits
  best_score_only = FALSE      # If best_score_only = FALSE, it will return ESet, its corresponding
                               # best score, and input_score for over 7 feature searches
)

# Now we can fetch the ESet and feature that corresponded to the best score over the top N search
revealer_topn_best_meta <- topn_best(revealer_topn_l)

# Visualize best meta-features result
meta_plot(topn_best_list = revealer_topn_best_meta)

```

## 4. Custom-Defined scoring method

See `?custom_genescore_mat` for more details

```{r custom.method}

# A customized function using ks-test function
customized_genescore_mat <- function(mat, input_score, alternative){
  result <- 1:nrow(mat) %>% 
    purrr::map_dfr(
      function(r){ 
        feature = mat[r,];
        x = input_score[which(feature==1)]; y = input_score[which(feature==0)];
        res <- ks.test(x, y, alternative=alternative)
        return(data.frame(score=res$statistic, p_value=res$p.value))
   })
}

# Search for best features using a custom-defined function
custom_topn_l <- CaDrA::topn_eval(
  ES = ES,
  input_score = input_score,
  custom_function = customized_genescore_mat,       # Use customized function
  custom_parameters = list(alternative = "less"),   # Additional parameters to pass to custom_function (i.e. other than ES and input_score)
  metric = "stat",                                  # Use pval to score features
  top_N = 7,                                        # Evaluate top 7 starting points for the search
  max_size = 10,                                    # Set max_size = 10 to account for 10 left-skewed features in sim.ES
  do_plot = FALSE,                                  # We will plot it AFTER finding the best hits
  best_score_only = FALSE                           # If best_score_only = FALSE, it will return ESet, its corresponding
                                                    # best score, and input_score for over 7 feature searches
)

# Now we can fetch the ESet and feature that corresponded to the best score over the top N search
custom_topn_best_meta <- topn_best(custom_topn_l)

# Visualize best meta-features result
meta_plot(topn_best_list = custom_topn_best_meta)

```

